### 2020-04-01 Model Data Sheets and Code/Zaca MCF Mortality Model Results.Rmd
added one more random forest trial. didnt add it to a binomial trial tho

mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
  select('drainage_assoication', 'ridge', 'mortality', 
         'prefire','mortality_qual', 'canopy_loss_success', 
         'canopy_loss_failure', 'postfire_canopy', 'postfire', 
         'Change_in_canopy_success', 'Change_in_canopy_failure', 
         'Change_Qualitative','high_mortality', 'area_scaled', 
         'dem_10m_zaca', 'flow_accumulation', 'sr', 'sr_new', 
         'rain', 'slope', 'tmax', 'tmean', 'tmin', 'TPI_five', 
         'TPI_queen', 'TWI', 'fri', 'dem_scaled', 'southness', "cwd", "vpd_jan", "vpd_aug" ) 
 #mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High")))
 # keep getting error (Outer names are only allowed for unnamed scalar atomic inputs) when order the levels, but the reordering works anyway (although it doesnt work with the ssame error in the binomial code. The code directly below works to reorder the levels without error)
mcf$mortality_qual  <- factor(mcf$mortality_qual, levels = c("None", "Low", "Moderate", "High"))

@@

Which variables contribute to mortality?

Literature review:

@@

>>>>

mcf <- read.csv('Zaca MCF variables for model.csv') %>% 
  select('drainage_association', 'ridge', 'mortality', 
         'prefire','mortality_qual', 'canopy_loss_success', 
         'canopy_loss_failure', 'postfire_canopy', 'postfire', 
         'Change_in_canopy_success', 'Change_in_canopy_failure', 
         'Change_Qualitative','high_mortality', 'area_scaled', 
         'dem_10m_zaca', 'flow_accumulation', 'sr', 'sr_new', 
         'rain', 'slope', 'tmax', 'tmean', 'tmin', 'TPI_five', 
         'TPI_queen', 'TWI', 'fri', 'dem_scaled', 'southness', "cwd", "vpd_jan", "vpd_aug", "tc_mortality" ) 
 #mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("None", "Low", "Moderate", "High")))
 # keep getting error (Outer names are only allowed for unnamed scalar atomic inputs) when order the levels, but the reordering works anyway (although it doesnt work with the ssame error in the binomial code. The code directly below works to reorder the levels without error)
mcf$mortality_qual  <- factor(mcf$mortality_qual, levels = c("None", "Low", "Moderate", "High"))

@@

@@

mcf_mod <- mcf %>% 
  mutate(tc_mortality = case_when(
    tc_mortality == 1 ~ "tc", #tc = type converted
    tc_mortality == 0 ~ "not_tc"
  )) %>% 
  mutate(tc_mortality = as.factor(tc_mortality))

# set sample size 
nsample = 75

#random forest code---------------------------------------------------
for (i in 1:50)
{

  # divide data into equal # of observations --------------------
  mcf_high <- mcf_mod[which(mcf_mod$tc_mortality == "tc"),]
  mcf_low <- mcf_mod[which(mcf_mod$tc_mortality == "not_tc"),]

  # pull random observations from the high class and separate
  x_high <- sample(1:nrow(mcf_high), nsample , replace = FALSE)
  polygon_data_high_train <- mcf_high[x_high,]
  polygon_data_high_test <- mcf_high[-x_high,]
  # pull random observations from the low class and separate
  x_low <- sample(1:nrow(mcf_low), nsample , replace = FALSE)
  polygon_data_low_train <- mcf_low[x_low,]
  polygon_data_low_test <- mcf_low[-x_low,]

  # combine all random observations
  train <- do.call("rbind", list(polygon_data_low_train, polygon_data_high_train))
  # recombine classes that were not pulled
  test <- do.call("rbind", list(polygon_data_low_test, polygon_data_high_test))

  #random forest model
  rfm<-randomForest(tc_mortality~ rain + tmax + TPI_five + dem_10m_zaca + flow_accumulation + sr + slope + TWI + cwd + vpd_aug + ridge + drainage_association, data = train, ntree=500, importance=T, nscale=0)

  #test data accuracy
  tabb<- table(pp, test$tc_mortality)
  acc<-1-sum(diag(tabb))/sum(tabb) # % missclassified 
  test_accuracy <-(1-acc)*100 #model accuracy

  # lists to save model outputs
  test_accuracy_list <- append(test_accuracy_list, test_accuracy)
  write.table(model_importance, file="rfm model tc importance.csv", sep=",",col.names=NA, append=T) 
  write.table(tabb, file = "mcf tc test accuracy table.csv", sep=",",col.names=NA, append=T)
}

model_importance_tidy <- read_csv(here("Model Data Sheets and Code", "rfm model tc importance.csv")) %>% 
  clean_names() %>% 
  filter(not_tc != "not_tc") %>% 
  rename (variables = x1) %>% 
  mutate (variables = str_to_lower(variables)) %>% 
  mutate (not_tc = as.numeric(not_tc),
          tc = as.numeric(tc),
          mean_decrease_accuracy = as.numeric(mean_decrease_accuracy),
          mean_decrease_gini = as.numeric(mean_decrease_gini)
          )

# model importance summary statistics
model_importance_summary <- model_importance_tidy %>% 
  group_by(variables) %>% 
  summarize (not_tc_mean = mean(not_tc),
             tc_mean = mean(tc),
             mean_dec_accuracy_mean = mean(mean_decrease_accuracy),
             mean_decrease_gini_mean = mean(mean_decrease_gini)
  )
file.remove(here::here("Model Data Sheets and Code", "rfm model tc importance.csv"))
# average class accuracy ---------------------------------------------------------

not_tc <- read_csv("mcf tc test accuracy table.csv") 
tc <- read_csv("mcf tc test accuracy table.csv")

not_tc_acc <- not_tc %>% 
  select (X1, not_tc) %>% 
  drop_na() %>% 
  filter (X1 == "not_tc") %>% 
  mutate (not_tc = as.numeric(not_tc),
          class_acc = not_tc / (636-nsample)) %>% 
  summarise(mean = mean(class_acc))

tc_acc <- tc %>% 
  select (X1, tc) %>% 
  drop_na() %>% 
  filter (X1 == "tc") %>%
  mutate (tc = as.numeric(tc),
          class_acc = tc / (99-nsample)) %>% 
  summarise(mean = mean(class_acc))

file.remove(here::here("Model Data Sheets and Code", "mcf tc test accuracy table.csv"))

```
 Didnt get to ALE plots. But model importance shows that slope and elevation are highly important for the type converted class. Not surprised because a lot of stands with the highest mortality were at the top of Big pine mountain which was at the highest elevation in our study area. This could indicate that this is trend is specific to this fire situation and does not necessarily predict that stands with high mortality are highly likely to occur in high elevation zones.
